Augmenting the DonorsChoose.org Corpus for Meta-Learning

### 2019-02-12 doc/paper/main.tex
paper/main.tex: Incorporate general peer-review
Tweak the introduction to hopefully make it more clear what is being
addressed in this research. Criticize other datasets more and outline
why the RMSE was not chosen. Further, better describe the process of
calculating the mean recall-position in a Top-N set.

Furthermore, drop supervisors except for Andrew, i.e. drop Joeran and
Akiko from the list of authors. Change the repository link to point to
GitHub:Edenhofer instead of GitHub:BeelGroup.

Additionally fix some typos and marginally shrink the space between
figures and text.

	\usepackage[font=small,labelfont=bf]{caption}  % Customize caption aesthetics
	\usepackage[font=small]{subcaption}  % Customize sub-captions

@@

	\makeatletter
	% Insert a float barrier before subsections just like the package 'placeins' does for sections
	%\AtBeginDocument{%
	%	\expandafter\renewcommand\expandafter\subsection\expandafter{%
	%		\expandafter\@fb@secFB\subsection
	%	}%
	%}
	\makeatother
	\setlength{\intextsep}{1.1em}  % Reduce excessive spacing between float and text
	\setlength{\textfloatsep}{1.1em}  % Reduce excessive spacing between caption and text

@@

\titlerunning{Augmenting the DonorsChoose.org Corpus for Meta-Learning}  % Abbreviated title
\title{Augmenting the DonorsChoose.org Corpus for Meta-Learning%
	\thanks{This publication emanated from research conducted with the financial support of Science Foundation Ireland (SFI) under Grant Number 13/RC/2106.}
}
\author{Gordian~Edenhofer\inst{1,2} \and Joeran~Beel\inst{1,2} \and Andrew~Collins\inst{1} \and Akiko~Aizawa\inst{2}}
\authorrunning{G. Edenhofer et al.}  % Abbreviated author field
\institute{
	Trinity College Dublin, School of Computer Science \& Statistics, ADAPT Centre, Ireland\\ \and
	National Institute of Informatics Tokyo, Digital Content and Media Sciences Division, Japan\\ \email{beelj@tcd.ie, collinsa@tcd.ie, aizawa@nii.ac.jp}
}

@@

\begin{abstract}
The DonorsChoose.org dataset of past donations provides a big and feature-rich corpus of users and items. The dataset matches donors to projects in which they might be interested in and hence is intrinsically about recommendations. Due to the availability of detailed item-, user- and transaction-features, this corpus represents a suitable candidate for meta-learning approaches to be tested. This study aims at providing an augmented corpus for further recommender systems studies to test and evaluate meta-learning approaches. In the augmentation, metadata of collaborative and content-based filtering techniques is amended to the corpus. It is further extended with aggregated statistics for users and transactions and an exemplary meta-learning experiment. The performance in the learning subsystem is measured via the recall of recommended items in a Top-N test set. The augmented dataset and the source code are released into the public domain at GitHub:BeelGroup/Augmented-DonorsChoose.org-Dataset.

@@

\section{Introduction}
Meta-Learning is the process of applying machine learning algorithms on metadata as is generated by a subsystem composed of machine learners. The goal of meta-learning systems is to improve existing approaches by combining the strengths of several single machine learning systems. In recommender systems it is an emerging field of interest as meta-learning promises to better sort through item, user and transaction information as to retrieve the best possible match of item and user.

One challenge for researchers interested in meta-learning is that suitable datasets are time-consuming and cumbersome to construct. Creating a system of individual machine learning experiments and amending their metadata to the original corpus takes time and resources away from addressing more important challenges in meta-learning for recommender systems. Multiple examples of manual and repetitious data augmentation may be found in the scientific literature, for instance in~\cite{CUNHA2018128,DBLP:journals/corr/abs-1805-12118,Ekstrand:2012:RFP:2365952.2366002}.

Our objective is to address the data augmentation part in a comprehensive way as to allow researchers to focus on the actual meta-learning evaluation. We are presenting a public domain dataset which to our best knowledge has not been used for meta-learning before in the scientific community, yet due to its feature-rich, vast corpus is a suitable candidate. The raw dataset is augmented with metadata of several machine learning algorithms from a variety of fields (cf.~\cite{DBLP:journals/corr/abs-1805-12118}). It is further amended with aggregated statistics about users and items and metadata of an exemplary experiment.

@@

\begin{table}[ht]
	\centering
	\caption{Illustrative example of the overall design of the augmented transaction table with amended learning subsystem performance scores, statistics and meta-learner information.}
	\label{tab:IllustrativeTable}
	\includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{{{../res/Illustrative example of the overall design of the augmented transaction table}}}
\end{table}

\subsection{Dataset Preparation}
To prepare the dataset for evaluations, duplicate interactions are merged as the recommender systems shall not recommend items to which the user has already donated to. Internal stop words in description strings of items are stripped. Furthermore, transactions containing no information about the user's location are dropped. Thereby enabling the consistent usage of a user’s location information in new approaches.

@@

A classical 5-fold cross-validation on the users is applied for training the collaborative filters. Using the recall in a Top-N test set, a score of how good an algorithm is able to recommend an item is calculated. The algorithm is provided with all the items with which the user from the test set has interacted with except for one and is requested to create a ranking of $100$ random items with which the user has not interacted with plus the one with which an interaction has taken place. An ordering is assumed to be implied by the magnitudes of the values in the reconstructed matrix of the decomposition respectively the distance to the neighbor. The position of the interacted item in the set of $101$ items is then used as score for the user-item transaction.

The validation process of the content-based approaches uses leave-one-out. For each transaction a profile of the user is constructed which excludes this transaction. Afterwards, the algorithm is requested to create a ranking of $100$ items with which the user has not interacted with and the one with which a transaction has taken place. The ranking is performed using the cosine similarity of an item’s vector and the user's profile.

The choice of evaluation method ensures that it can be applied to a vast set of different algorithms as long as a ranking of items can be produced and therefore allow for a high variety within the set of algorithms in the learning subsystem. It further ensures that not just the ability to decompose the data is measured but the ability to recommend new items to users~\cite{Cremonesi:2010:PRA:1864708.1864721}. Hereby, the learning subsystem is kept extensible while still providing useful metadata to meta-learning algorithms.

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{{{../res/Learning subsystem - Position in Top-N test set}}}
	\caption{Statistical features of the distributions of the recall-positions of the learning subsystem. The dashed green lines indicate the mean, the straight orange lines the median. The percentage below each algorithm indicates how often it is the best possible algorithm. To resolve a draw deterministically, the ``best'' algorithm is decided in alphabetic order.}
	\label{fig:MetaLearnerStatistics}
\end{figure}

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{{{../res/Meta-learner Performance - Average position in Top-N test set for various meta-learner algorithms with augmented learning subsystem filtering techniques}}}
	\caption{Performance of the meta-learner system addressing the ASP. The dashed orange line represents the overall best algorithm, i.e., TF-IDF.}
	\label{fig:MetaLearnerPerformance}
\end{figure}

>>>>

	\usepackage[inline]{enumitem}  % Inline enumeration capability
	\usepackage[font=small,labelfont=bf]{caption}  % Customize caption aesthetics
	\usepackage[font=small]{subcaption}  % Customize sub-captions

@@

	\makeatletter
	% Insert a float barrier before subsections just like the package 'placeins' does for sections
	%\AtBeginDocument{%
	%	\expandafter\renewcommand\expandafter\subsection\expandafter{%
	%		\expandafter\@fb@secFB\subsection
	%	}%
	%}
	\makeatother
	\setlength{\intextsep}{1em}  % Reduce excessive spacing between float and text
	\setlength{\textfloatsep}{1em}  % Reduce excessive spacing between caption and text

@@

\titlerunning{Augmenting the DonorsChoose.org Corpus for Meta-Learning}  % Abbreviated title
\title{Augmenting the DonorsChoose.org Corpus for Meta-Learning}
\author{Gordian~Edenhofer\inst{1,2} \and Andrew~Collins\inst{1}}
\authorrunning{G. Edenhofer et al.}  % Abbreviated author field
\institute{
	Trinity College Dublin, School of Computer Science \& Statistics, ADAPT Centre, Ireland\\ \and
	National Institute of Informatics Tokyo, Digital Content and Media Sciences Division, Japan\\ \email{G.Edenhofer@campus.lmu.de, collinsa@tcd.ie}
}

@@

\begin{abstract}
The DonorsChoose.org dataset of past donations provides a big and feature-rich corpus of users and items. The dataset matches donors to projects in which they might be interested in and hence is intrinsically about recommendations. Due to the availability of detailed item-, user- and transaction-features, this corpus represents a suitable candidate for meta-learning approaches to be tested. This study aims at providing an augmented corpus for further recommender systems studies to test and evaluate meta-learning approaches. In the augmentation, metadata of collaborative and content-based filtering techniques is amended to the corpus. It is further extended with aggregated statistics for users and transactions and an exemplary meta-learning experiment. The performance in the learning subsystem is measured via the recall of recommended items in a Top-N test set. The augmented dataset and the source code are released into the public domain at GitHub:Edenhofer/Augmented-DonorsChoose.org-Dataset.

@@

\section{Introduction}
Meta-Learning is the process of applying machine learning algorithms on metadata as is generated by a subsystem composed of machine learners. The goal of meta-learning systems is to improve existing approaches by combining the strengths of several single machine learning systems in an ensemble. In recommender systems it is an emerging field of interest as meta-learning promises to better sort through item, user and transaction information as to retrieve the best possible match of item and user.

One challenge for researchers interested in meta-learning is that suitable datasets are time-consuming and cumbersome to construct. Creating a system of individual machine learning experiments and amending their metadata to the original corpus takes time and resources away from working with the actual meta-learners. Multiple examples of manual and repetitious data augmentation may be found in the scientific literature, for instance in~\cite{CUNHA2018128,DBLP:journals/corr/abs-1805-12118,Ekstrand:2012:RFP:2365952.2366002}. Furthermore, a noteworthy shortcoming of the datasets constructed in the aforementioned publications is the usage of a metric which limits the choice of algorithms. The RMSE of the predicted and actual score is a popular metric but, e.g., can not be computed in a straight-forward way for content-based filtering algorithms.

Our objective is to address the data augmentation part in a comprehensive way as to allow researchers to focus on the actual meta-learning evaluation. We are presenting a public domain dataset which to our best knowledge has not been used for meta-learning before in the scientific community, yet due to its feature-rich, vast corpus is a suitable candidate. The raw dataset is augmented with metadata of several machine learning algorithms from a variety of fields (cf.~\cite{DBLP:journals/corr/abs-1805-12118}). In place of a metric, the recall in Top-N set is chosen as it can be computed for a wide range of algorithms. It is further amended with aggregated statistics about users and items and metadata of an exemplary experiment.

@@

\begin{table}[ht]
	\centering
	\caption{Illustrative example of the overall design of the augmented transaction table with amended learning subsystem performance scores, statistics and meta-learner information.}%
	\label{tab:IllustrativeTable}
	\includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{{{../res/Illustrative example of the overall design of the augmented transaction table}}}
\end{table}

\subsection{Dataset Preparation}
To prepare the dataset for evaluations, duplicate interactions are merged as the recommender systems shall not recommend items to which the user has already donated to. Internal stop words in description strings of items are stripped. Furthermore, transactions containing no information about the user's location are dropped. Thereby enabling the consistent usage of a user's location information in new approaches.

@@

The recall in a Top-N test set is used as score of how good an algorithm is able to recommend an item.
\begin{enumerate*}[label= (\arabic*)]
	\item For each user, first $100$ items with which the user has not interacted with are selected.
	\item The algorithm is then requested to rank an item with which the user has interacted with but which was not used for building the user's profile and the $100$ randomly sampled items. It is assumed that the random $100$ items are of low interest to the user.
	\item The position of the item with which the user has interacted with in the ranked test set may now be taken as the recall in a Top-N test set.
\end{enumerate*}

A classical 5-fold cross-validation on the users is applied for training the collaborative filters.  An ordering is assumed to be implied by the magnitudes of the values in the reconstructed matrix of the decomposition respectively the distance to the neighbor. The validation process of the content-based approaches uses leave-one-out. The ranking is performed using the cosine similarity of an item's vector and the user's profile.

The choice of evaluation method ensures that it can be applied to a vast set of different algorithms as long as a ranking of items can be produced. Therefore, it allow for a high variety within the set of algorithms in the learning subsystem in contrast to, e.g., RMSE. It further ensures that not just the ability to decompose the data is measured but the ability to recommend new items to users~\cite{Cremonesi:2010:PRA:1864708.1864721}. Hereby, the learning subsystem is kept extensible while still providing useful metadata to meta-learning algorithms.

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{{{../res/Learning subsystem - Position in Top-N test set}}}
	\caption{Statistical features of the distributions of the recall-positions of the learning subsystem. The dashed green lines indicate the mean, the straight orange lines the median. The percentage below each algorithm indicates how often it is the best possible algorithm. To resolve a draw deterministically, the ``best'' algorithm is decided in alphabetic order.}%
	\label{fig:MetaLearnerStatistics}
\end{figure}

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{{{../res/Meta-learner Performance - Average position in Top-N test set for various meta-learner algorithms with augmented learning subsystem filtering techniques}}}
	\caption{Performance of the meta-learner system addressing the ASP. The dashed orange line represents the overall best algorithm, i.e., TF-IDF.}%
	\label{fig:MetaLearnerPerformance}
\end{figure}
